---
title: "Reproducible Research Peer Grading Assignment.Rmd"
author: "shas"
date: "01-08-2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading and preprocessing the data:
```{r, echo = TRUE, warning=FALSE, message = FALSE}
library(dplyr)
library(ggplot2)
library(skimr)
library(Hmisc)
library(lubridate)

Sys.setlocale("LC_TIME", "English")
```

```{r, echo = TRUE, warning=FALSE, message = FALSE}
data <- read.csv("activity.csv", header = TRUE, na.strings=c("",".","NA"))
head(data)
str(data)
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE, warning=FALSE, message = FALSE}
total_steps <- data %>% 
    group_by(date) %>% 
    summarise(daily_steps = sum(steps))

head(total_steps)
```

Histogram of total number of steps taken per day:

```{r, echo = TRUE, warning=FALSE, message = FALSE}

ggplot(data = total_steps, aes(daily_steps)) +
    geom_histogram(col="darkblue", fill="lightblue") +
    ggtitle("Histogram of steps per day") +
    xlab("Total Number of Steps Taken Per Day") +
    ylab("Frequency") +
    theme_bw()
```

## Mean and median number of steps taken each day:

```{r, echo = TRUE, warning=FALSE, message = FALSE}
mean(total_steps$daily_steps, na.rm = TRUE)
median(total_steps$daily_steps, na.rm = TRUE)
```

## The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo = TRUE, warning=FALSE, message = FALSE}
average_interval_steps <- data %>% 
    group_by(interval) %>% 
    summarise(average_steps = mean(steps, na.rm = TRUE))
```

```{r, echo = TRUE, warning=FALSE, message = FALSE}
ggplot(data = average_interval_steps, aes(x = interval, y = average_steps)) +
    geom_line(color = "blue", size = 1) +
    ggtitle("Average steps per time interval") +
    xlab("Time") +
    ylab("Steps") +
    theme_bw()
```

## 5-minute interval that contains the maximum number of steps is the 835th interval.

```{r, echo = TRUE, warning=FALSE, message = FALSE}
average_interval_steps[which.max(average_interval_steps$average_steps),]$interval
```

## Code to describe and show a strategy for imputing missing data:

```{r, echo = TRUE, warning=FALSE, message = FALSE}
str(data)
skim_without_charts(data)
```

"Steps" variable has 2304 missing values

I will replace the missing values in "steps" variable with the average values of it. First I created a new variable "activity" that is the same as the original data, then filled the missing values with the average values in "steps" variable, which is 37.3826

```{r, echo = TRUE, warning=FALSE, message = FALSE}
activity <- data 

activity$steps <- impute(data$steps, fun = mean)
```

## Histogram of the total number of steps taken each day after missing values are imputed:

```{r, echo = TRUE, warning=FALSE, message = FALSE}
imputed_total_steps <- activity %>% 
    group_by(date) %>% 
    summarise(imputed_total_steps = sum(steps))
```

```{r, echo = TRUE, warning=FALSE, message = FALSE}
ggplot(data = imputed_total_steps , aes(imputed_total_steps)) +
    geom_histogram(col="darkblue", fill="lightblue") +
    xlab("Total Number of Steps Taken Per Day") +
    ylab("Frequency") +
    theme_bw()
```

```{r, echo = TRUE, warning=FALSE, message = FALSE}
mean(imputed_total_steps$imputed_total_steps, na.rm = TRUE)
median(imputed_total_steps$imputed_total_steps, na.rm = TRUE)
```

After filling the missing values, unsurprisingly, the mean value of steps didn't change because I filled the missing values with the average value. But the median value changed from 10765 to 10766.19.

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends:

```{r, echo = TRUE, warning=FALSE, message = FALSE}
day_of_week <- activity %>%
    mutate(
        date = ymd(date),
        weekday_or_weekend = case_when(wday(date) %in% 2:6 ~ "Weekday",
                                       wday(date) %in% c(1,7) ~ "Weekend")
    ) %>%
    group_by(interval, weekday_or_weekend) %>%
    summarise(
        steps = mean(steps)
    )
```

```{r, echo = TRUE, warning=FALSE, message = FALSE}
head(day_of_week)
```

```{r, echo = TRUE, warning=FALSE, message = FALSE}
ggplot(day_of_week, aes(interval, steps)) + 
    geom_line(color = "blue", size = 1) + 
    facet_wrap(~weekday_or_weekend, nrow = 2, ncol = 1) +
    xlab("5-Minute intervals") + 
    ylab("Average number of steps") +
    theme_bw()
```

## Are there differences in activity patterns between weekdays and weekends?

On weekdays, most of the steps are done between 500-1000th intervals.
On weekends, most of the steps are done between 1000-1750th intervals.






